Estimating the Effect of Urban Green Spaces on Residential House Prices in Amsterdam

### 2023-04-14 Paper/Term-Project---Grings,-Nauta,-Nibbering.Rmd
update

# 2. Methods
## 2.1 Study Area
The geographical scope in this study concentrated on the city of Amsterdam, as presented in Figure 1. The city is situated in the Western part of the Netherlands and home to almost 880,000 residents [27], occupying a space of 219.3 km2  [28]. Surrounding the residential areas, the vast majority of land is occupied with vegetation [29]. Moreover, the municipality has progressive urban greening programs in place, stimulating and obligating residents and city districts to participate in greening initiatives [12]. As a result, these policies aimed at expanding urban green spaces may have the potential to increase nearby housing premiums.

```{=latex}
\begin{figure}[]
  \centering
  \includegraphics[width=\linewidth]{figures/study_area.png}
  \caption{Overview of the study area.}
\end{figure}
```

@@

```{=latex}
\begin{table*}[]
\centering
\caption{Descriptive statistics of property characteristics, public amenities and greenness measures of Amsterdam real-estate.}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Independent Variable}      & \textbf{Mean} & \textbf{SD} & \textbf{Description}                        \\ \midrule
\multicolumn{4}{c}{\textbf{Summary Statistics (N = 1,719)}}                                                    \\ \midrule
\textbf{Property Characteristics$^1$}  &               &             &                                             \\
Living Area (m2) [38]                   & 105.7         & 71.0        & Size of livable area of the property        \\
Number of Bedrooms [38, 39]                & 2.5           & 1.4         & Number of bedrooms of the property          \\
Number of Bathrooms [23]               & 1.1           & 0.7         & Number of bathrooms of the property         \\
House Age (years) [23, 38]            & 76.9          & 63.2        & Age of the property in 2023                 \\
\textbf{Design$^2$}                    &               &             &                                             \\
NDVI$_{100}$ [36, 39, 40]                & 0.200         & 0.073       & Average NDVI at 100 m around property       \\
NDVI$_{300}$ [36, 39, 40]                & 0.212         & 0.069       & Average NDVI at 300 m around property       \\
NDVI$_{500}$ [36, 39, 40]               & 0.216         & 0.067       & Average NDVI at 500 m around property       \\
\textbf{Transit Distance$^{3,4,5}$} &               &             &                                             \\
Distance to Train Station (m) [22]     & 1703.7        & 958.2       & Euclidean distance to nearest train station \\
Distance to Tram Station (m) [22]      & 689.1         & 1024.7      & Euclidean distance to nearest tram station  \\
Distance to Metro Station (m) [23]      & 1303.2        & 951.8       & Euclidean distance to nearest metro station \\
\textbf{Destination Accessibility$^4$} &               &             &                                             \\
Distance to City Center (m) [38]       & 3770.5        & 2288.1      & Euclidean distance to city center           \\
Distance to Business District (m) [38] & 4659.2 & 2228.3 & Euclidean distance to business district \\
Distance to School (m) [23]            & 268.3         & 183.8       & Euclidean distance to nearest school        \\
Distance to Convenience Stores (m) [38] & 266.7         & 206.9       & Euclidean distance to nearest store         \\ \bottomrule
\multicolumn{4}{l}{Sources: $^1$Funda.nl, $^2$Sentinel-2, $^3$ArcGIS, $^4$OpenStreetMap.org, and $^5$Amsterdam Municipality} 
\end{tabular} 
\end{table*}
```

@@

Results of the local regression model revealed that urban green spaces were significantly associated with residential house prices at a 300m spatial scale (p < 0.05). Here, property premiums per m2 ranged from -82,657.5 to 125,978.4 euros for each unit increase in the vegetation index. This result aligns with previous studies [23, 36, 39, 40]. One explanation for this wide range might be related to the direct-use value placed on these areas [23]. In other words, particular green areas might attract more residents to benefit from the perceived value of greenness, prompting more people to purchase houses nearby [44]. Hence, highlighting the spatial dependent and non-stationary nature of the association of interest, as illustrated in Appendix II. Finally, the findings of this study provide empirical evidence of the capitalisation of greenness characteristics in residential real-estate prices. \linebreak

@@

# 4. Conclusion
This study aimed to examine the economic value of urban green spaces in Amsterdam, as reflected in the real-estate prices of 2023. Here, property characteristics, public amenities and greenness measurements were obtained from a variety of sources. In line with topical literature, independent variables were identified and constructed to either adjust or explain the economic value of urban green spaces. In doing so, the local hedonic pricing model revealed that urban green spaces provided significant economic value. However, this association was not consistent across all locations, as the property premiums per m2 ranged from -82,657.5 to 125,978.4 euros for each unit increase in the vegetation index. The main theoretical explanation for this range might be related to the direct-use value placed on these areas. As a result, the outcomes of this study provide empirical evidence for the capitalisation of greenness characteristics in residential real-estate. \linebreak

@@

```{=latex}
\begin{table*}[ht]
\centering
\caption{Coefficients and statistical significance of the Ordinary Least Squares and Geographically Weighted Regression models.}
\begin{tabular}{@{}lllll@{}}
\toprule
\textbf{Independent Variable} &
  \textbf{Coef.} &
  \textbf{Min. Coef.} &
  \textbf{Median Coef.} &
  \textbf{Max. Coef.} \\ \midrule
                                     & OLS$^1$ &           & GWR$^{1,2}$ &           \\ \midrule
\textbf{Property Characteristics}    &            &           &         &           \\
Living Area (m2)                     & 4.7**      & -45.7     & -5.3    & 28.4      \\
Number of Bedrooms                   & -329.6**   & -1384.2   & -136.0  & 957.8     \\
Number of Bathrooms                  & 318.3**    & -1206.8   & 248.3   & 2,385.9   \\
House Age (years)                    & 1.3**      & -121.1    & -0.6    & 74.1      \\
\textbf{Design}                      &            &           &         &           \\
NDVI$_{300}$                              & -1,674.1** & -82,657.5 & 262.5   & 125,978.4 \\
\textbf{Transit Distance}            &            &           &         &           \\
Distance to Train Station (m)        & -0.08      & -79.7     & 1.0     & 297.5     \\
Distance to Tram Station (m)         & 0.4**      & -38.5     & 0.2     & 45.8      \\
Distance to Metro Station (m)        & 0.5**      & -66.0     & 0.5     & 39.1      \\
\textbf{Destination Accessibility}   &            &           &         &           \\
Distance to City Center (m)          & -0.5**     & -495.3    & -1.2    & 124.4     \\
Distance to Business District   (m)  & -0.5**     & -145.6    & -0.02   & 162.1     \\
Distance to School (m)               & 0.3        & -21.5     & -0.6    & 9.4       \\
Distance to Convenience Stores   (m) & 1.2**      & -15.4     & 0.2     & 18.1      \\ \midrule
Akaike Information Criterion   (AIC) & 30,487     &           & 29,720  &           \\ \midrule
\multicolumn{5}{l}{** Significant at p \textless 0.05}                              \\
\multicolumn{5}{l}{$^1$ Statistical significance levels are only reported for the global OLS model,} \\
\multicolumn{5}{l}{as p-values could not be easily presented for GWR due to its local approach.} \\
\multicolumn{5}{l}{$^2$ Bandwidth of adjusted kernel corresponds to 67.}              
\end{tabular}
\end{table*}
```

>>>>

# 2. Methods
## 2.1 Study Area
The geographical scope in this study concentrated on the city of Amsterdam, as presented in Figure 1. The city is situated in the Western part of the Netherlands and home to almost 880,000 residents [27], occupying a space of 219.3 km$^2$  [28]. Surrounding the residential areas, the vast majority of land is occupied with vegetation [29]. Moreover, the municipality has progressive urban greening programs in place, stimulating and obligating residents and city districts to participate in greening initiatives [12]. As a result, these policies aimed at expanding urban green spaces may have the potential to increase nearby housing premiums.

```{=latex}
\begin{figure}[h]
  \centering
  \includegraphics[width=.8\linewidth]{figures/study_area.png}
  \caption{Overview of the study area.}
\end{figure}
```

@@

```{=latex}
\begin{table*}[]
\centering
\caption{Descriptive statistics of property characteristics, public amenities and greenness measures of Amsterdam real-estate.}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Independent Variable}      & \textbf{Mean} & \textbf{SD} & \textbf{Description}                        \\ \midrule
\multicolumn{4}{c}{\textbf{Summary Statistics (N = 1,719)}}                                                    \\ \midrule
\textbf{Property Characteristics$^1$}  &               &             &                                             \\
Living Area (m$^2$) [38]                   & 105.7         & 71.0        & Size of livable area of the property        \\
Number of Bedrooms [38, 39]                & 2.5           & 1.4         & Number of bedrooms of the property          \\
Number of Bathrooms [23]               & 1.1           & 0.7         & Number of bathrooms of the property         \\
House Age (years) [23, 38]            & 76.9          & 63.2        & Age of the property in 2023                 \\
\textbf{Design$^2$}                    &               &             &                                             \\
NDVI$_{100}$ [36, 39, 40]                & 0.200         & 0.073       & Average NDVI at 100 m around property       \\
NDVI$_{300}$ [36, 39, 40]                & 0.212         & 0.069       & Average NDVI at 300 m around property       \\
NDVI$_{500}$ [36, 39, 40]               & 0.216         & 0.067       & Average NDVI at 500 m around property       \\
\textbf{Transit Distance$^{3,4,5}$} &               &             &                                             \\
Distance to Train Station (m) [22]     & 1703.7        & 958.2       & Euclidean distance to nearest train station \\
Distance to Tram Station (m) [22]      & 689.1         & 1024.7      & Euclidean distance to nearest tram station  \\
Distance to Metro Station (m) [23]      & 1303.2        & 951.8       & Euclidean distance to nearest metro station \\
\textbf{Destination Accessibility$^4$} &               &             &                                             \\
Distance to City Center (m) [38]       & 3770.5        & 2288.1      & Euclidean distance to city center           \\
Distance to Business District (m) [38] & 4659.2 & 2228.3 & Euclidean distance to business district \\
Distance to School (m) [23]            & 268.3         & 183.8       & Euclidean distance to nearest school        \\
Distance to Convenience Stores (m) [38] & 266.7         & 206.9       & Euclidean distance to nearest store         \\ \bottomrule
\multicolumn{4}{l}{Sources: $^1$Funda.nl, $^2$Sentinel-2, $^3$ArcGIS, $^4$OpenStreetMap.org, and $^5$Amsterdam Municipality} 
\end{tabular} 
\end{table*}
```

@@

Results of the local regression model revealed that urban green spaces were significantly associated with residential house prices at a 300m spatial scale (p < 0.05). Here, property premiums per m$^2$ ranged from -82,657.5 to 125,978.4 euros for each unit increase in the vegetation index. This result aligns with previous studies [23, 36, 39, 40]. One explanation for this wide range might be related to the direct-use value placed on these areas [23]. In other words, particular green areas might attract more residents to benefit from the perceived value of greenness, prompting more people to purchase houses nearby [44]. Hence, highlighting the spatial dependent and non-stationary nature of the association of interest, as illustrated in Appendix II. Finally, the findings of this study provide empirical evidence of the capitalisation of greenness characteristics in residential real-estate prices. \linebreak

@@

# 4. Conclusion
This study aimed to examine the economic value of urban green spaces in Amsterdam, as reflected in the real-estate prices of 2023. Here, property characteristics, public amenities and greenness measurements were obtained from a variety of sources. In line with topical literature, independent variables were identified and constructed to either adjust or explain the economic value of urban green spaces. In doing so, the local hedonic pricing model revealed that urban green spaces provided significant economic value. However, this association was not consistent across all locations, as the property premiums per m$^2$ ranged from -82,657.5 to 125,978.4 euros for each unit increase in the vegetation index. The main theoretical explanation for this range might be related to the direct-use value placed on these areas. As a result, the outcomes of this study provide empirical evidence for the capitalisation of greenness characteristics in residential real-estate. \linebreak

@@

```{=latex}
\begin{table*}[ht]
\centering
\caption{Coefficients and statistical significance of the Ordinary Least Squares and Geographically Weighted Regression models.}
\begin{tabular}{@{}lllll@{}}
\toprule
\textbf{Independent Variable} &
  \textbf{Coef.} &
  \textbf{Min. Coef.} &
  \textbf{Median Coef.} &
  \textbf{Max. Coef.} \\ \midrule
                                     & OLS$^1$ &           & GWR$^{1,2}$ &           \\ \midrule
\textbf{Property Characteristics}    &            &           &         &           \\
Living Area (m$^2$)                     & 4.7**      & -45.7     & -5.3    & 28.4      \\
Number of Bedrooms                   & -329.6**   & -1384.2   & -136.0  & 957.8     \\
Number of Bathrooms                  & 318.3**    & -1206.8   & 248.3   & 2,385.9   \\
House Age (years)                    & 1.3**      & -121.1    & -0.6    & 74.1      \\
\textbf{Design}                      &            &           &         &           \\
NDVI$_{300}$                              & -1,674.1** & -82,657.5 & 262.5   & 125,978.4 \\
\textbf{Transit Distance}            &            &           &         &           \\
Distance to Train Station (m)        & -0.08      & -79.7     & 1.0     & 297.5     \\
Distance to Tram Station (m)         & 0.4**      & -38.5     & 0.2     & 45.8      \\
Distance to Metro Station (m)        & 0.5**      & -66.0     & 0.5     & 39.1      \\
\textbf{Destination Accessibility}   &            &           &         &           \\
Distance to City Center (m)          & -0.5**     & -495.3    & -1.2    & 124.4     \\
Distance to Business District   (m)  & -0.5**     & -145.6    & -0.02   & 162.1     \\
Distance to School (m)               & 0.3        & -21.5     & -0.6    & 9.4       \\
Distance to Convenience Stores   (m) & 1.2**      & -15.4     & 0.2     & 18.1      \\ \midrule
Akaike Information Criterion   (AIC) & 30,487     &           & 29,720  &           \\ \midrule
\multicolumn{5}{l}{** Significant at p \textless 0.05}                              \\
\multicolumn{5}{l}{$^1$ Statistical significance levels are only reported for the global OLS model,} \\
\multicolumn{5}{l}{as p-values could not be easily presented for GWR due to its local approach.} \\
\multicolumn{5}{l}{$^2$ Bandwidth of adjusted kernel corresponds to 67.}              
\end{tabular}
\end{table*}
```
